\batcave: Bayesian Analysis Tools for Context-Aware Variant Evaluation

### 2019-07-29 NAR_DRAFT.tex
check in

@@

We simulated realistic variant sites and allele frequencies using a branching process to simulate neutral evolution with no death.
Variants were selected from TCGA and PCAWG variant files(dates).
Whole genome (100X depth), and whole exome (500X depth) reads were simulated from the GRCH38 reference genome using VarSim \cite{Mu2015}, and aligned with BWA \cite{Li2009a}, both with default parameters.
Variants were spiked to create tumors with Bamsurgeon with default parameters \cite{Ewing2015a},
and called with MuTect 1.1.7 \cite{Cibulskis2013} with the following parameters:

@@

\subsection{Simulated data}
We generated six different tumor/normal pairs of two types, 3 each of 100X depth whole genomes and 500X whole exomes.
For each of the two types we simulated tumors with variants drawn from three different mutation profiles (Concentrated,Intermediate, and Diffuse) as described in Materials \& Methods.
These 6 simulated tumor/normal pairs were used to assess the effect of sequencing depth and mutation profile type on the results obtained from \batcave.
\bkmcomment{Should there be a longer discussion of what sorts of tumors the profiles represent, or leave that for the methods only?}

@@

>>>>

\bkmcomment{Introduce D earlier as its own section describing how we get it? Or expound here instead.}

@@

We simulated realistic variant sites and allele frequencies using a branching process to simulate neutral evolution with no death.
The simulated mutation rate was $\mu = 3\cdot10^{-6}$.
Variants were selected from TCGA and PCAWG variant files(dates).
Whole genome (100X depth), and whole exome (500X depth) reads were simulated from the GRCH38 reference genome using VarSim \cite{Mu2015}, and aligned with BWA \cite{Li2009a}, both with default parameters.
Variants were spiked to create tumors with Bamsurgeon with default parameters \cite{Ewing2015a},
and called with MuTect 1.1.7 \cite{Cibulskis2013} with the following parameters:

@@

\subsection{Simulated data}
We generated six different tumor/normal pairs of two types, 3 each of 100X depth whole genomes and 500X whole exomes.
For each of the two types we simulated tumors with variants drawn from three different mutation profiles (Concentrated,Intermediate, and Diffuse) as described in Materials \& Methods and Figure \ref{NAR-sigfig}.
These 6 simulated tumor/normal pairs were used to assess the effect of sequencing depth and mutation profile type on the results obtained from \batcave.
\bkmcomment{Should there be a longer discussion of what sorts of tumors the profiles represent, or leave that for the methods only?}

@@

The estimated effective mutation rate, $\mu/\beta$, is in the range of 3e-7 \bkmcomment{filling in today} for all simulated tumors.
We note that the effective mutation rate is lower than the simulated rate of 3e-7.
In this case the lower effective rate is not due to normal contamination of the tumor sample, but the fact that \texttt{Bamsurgeon} has restrictions -- such as sequencing depth and quality -- that prevent 100\% of simulated variants from being inserted. 
\bkmcomment{possible to quantify that this is very close to the actual insertion rate. Do it}.

We measured calibration performance using the Integrated Calibration Index described in Methods \citep{Austin2019}.
